---
title: "Final Project - Titanic Passengers"
author: "Iffah Mohamad Saufi"
output: html_notebook
---

#### 1) Research Question:
##### Does survivors mostly came from the first class during the Titanic incident?

#### 2) Data Sources
##### Primary Source: The data comes from Wikipedia. It was collected after the titanic incident since the survivors were also included in the data.

##### Second Source: From R package called vcdExtra.

##### There are 1309 cases and each of them represents the passenger of the Titanic ship. For this project, I am planning to use survived, class, age and gender variables. 

```{r}
# clean up R environment
rm(list = ls())

# load packages
library(tidyverse)
library(mosaic)
library(mdsr)
library(DataComputing)
library(vcdExtra)
library(dplyr)
```

```{r}
# Read Primary Data into R
Passenger <- read.csv("full.csv")

# inspect data
head(Passenger)
str(Passenger)
glimpse(Passenger)

# Read second Data Source
data("Titanicp")

# inspect data
head(Titanicp)
str(Titanicp)
glimpse(Titanicp)
```
```{r}
# changing the survive variable to numeric
Passenger$Survived <- as.numeric(Passenger$Survived)

# use loop to change Pclass and Sex to factor
for (i in c("Pclass", "Sex")){
  Passenger[,i] = as.factor(Passenger[,i])
}

str(Passenger)
```

```{r}
# Creating a new dataframe with some variables that will be used
Survivor <-
  Passenger %>%
  select("PassengerId", "Name", "Pclass", "Survived", "Sex", "Age", "Fare")

# eliminate incomplete subject
Survivor <- Survivor[!is.na(Survivor$Age),]
Survivor <- Survivor[!is.na(Survivor$Survived),]
head(Survivor)
```

```{r}
# spreading the Sex variable 
Gender <- 
  Survivor %>%
  group_by(Name, Sex) %>%
  spread(key = Sex, value = Survived)
head(Gender)

# counting the number of survivors between male and female
Female = sum(Gender$female, na.rm = TRUE)
Female
Male = sum(Gender$male, na.rm = TRUE)
Male
```
There are 197 female survivors and 93 male survivors.

```{r}
# Recall my reserarch question: Does survivors mostly came from the first class during the Titanic incident? 
# I want to see whether passenger's class is significant to the survivors.

RegressModel <- lm(Survived ~ Pclass, data = Survivor)
RegressModel$coefficients 
summary(RegressModel)
```
By using alpha = 0.5, it can be seen that the p-value for explanatory variable is less than alpha. Thus, Pclass is statistically significant to the survivors. It means that the survivors have relationship to which class they were from.

```{r}
# EDA for Titanic Passenger
# To see the relationship between Pclass, Age, Fare and Embarked
Passenger %>%
  ggplot(aes(x = Age, y = Fare, color = Pclass)) + 
  geom_smooth(method = "lm") + geom_point() + theme_bw()

# To see the relationship between Survivors, Age, Fare and Pclass
Survivor %>%
  ggplot(aes(x = Age, y = Fare, color = Pclass)) + 
  geom_point() + facet_grid(.~Survived) + theme_bw()

```
Bases on the first graph, I can see that fare for passengers from class three were indeed higher than other classes. Based on the second graph, I can see that from class two and three, there were more survivors. From class one, there were more unsurvivor passengers. 

To be more confident with my answer to my research question, I am adding one more graph by narrowing down the variables to see the relationship of survivors and classes as well as gender.

```{r}
# Plot to see the relationship of survivors and classes as well as gender

Survivor %>%
  ggplot(aes(Age, Survived, color = Pclass)) +
  stat_smooth(method = "glm", family = "binomial", formula = y~x, alpha = .2, aes(fill = Pclass)) +
  facet_grid(.~Sex) + theme_bw()
```
From the graph above, it can be seen that the survivors are more among female than male. For both genders, I can say that survivors mostly came from class one. 




