---
title: "Final Project - Titanic Passengers"
author: "Iffah Mohamad Saufi"
output: html_notebook
---

#### 1) Research Question:
##### Does survivors mostly came from the first class during the Titanic incident?

#### 2) Data Sources
##### Primary Source: The data comes from Wikipedia. It was collected after the titanic incident since the survivors were also included in the data.

##### Second Source: From R package called vcdExtra.

##### There are 1309 cases and each of them represents the passenger of the Titanic ship. For this project, I am planning to use survived, class, age and gender variables. 

```{r}
# clean up R environment
rm(list = ls())

# load packages
library(tidyverse)
library(mosaic)
library(mdsr)
library(DataComputing)
library(vcdExtra)
```

```{r}
# Read Primary Data into R
Passenger <- read.csv("full.csv")

# inspect data
head(Passenger)
str(Passenger)
glimpse(Passenger)

# Read second Data Source
data("Titanicp")

# inspect data
head(Titanicp)
str(Titanicp)
glimpse(Titanicp)
```

```{r}
# Informative plot for second data source
hist(Titanicp$age, xlab = "Age", main = "Age of Titanic Passengers")
```
From the bar graph above, we know that the average of the passenger is around 20-30 years old. 

```{r}
Titanic <- 
  Passenger %>%
  select(Pclass, Survived, Sex, Age, Name) %>%
  left_join(Titanicp %>% select(survived),
            by = c("Survived" = "survived"))
```

```{r}
# changing the variables
Titanic$Pclass <- as.factor(Titanic$Pclass)
Titanic$Survived <- as.numeric(Titanic$Survived)
Titanic$Sex <- as.factor(Titanic$Sex)
Titanic$Age <- as.numeric(Titanic$Age)

str(Titanic)
```

```{r}
# count the survivors
Titanic %>%
  group_by(Sex) %>%
  summarise(TotalSurvived = sum(Survived, na.rm = TRUE))
```

```{r}
# selecting variables that will be used
PassengerList <-
  Passenger %>%
  select("Survived", "Sex", "Pclass", "Age") 

# clean data
PassengerList <- PassengerList[!is.na(PassengerList$Age),] 
```

```{r}
# changing the variables
PassengerList$Pclass <- as.factor(PassengerList$Pclass)
PassengerList$Survived <- as.numeric(PassengerList$Survived)
str(PassengerList)
```

```{r}
# Plot to see the relationship of survivors and classes as well as gender

PassengerList %>%
  ggplot(aes(Age, Survived, color = Pclass)) +
  stat_smooth(method = "glm", family = "binomial", formula = y~x, alpha = .2, aes(fill = Pclass)) +
  facet_grid(.~Sex) +
  theme_bw()
```
From the graph above, it can be seen that the survivors are more among female than male. Between female passengers, the survivors mostly come from class two while between male passengers, survivors mostly come from class one. 




