---
title: "Final Project - Titanic Passengers"
author: "Iffah Mohamad Saufi"
output: html_notebook
---

#### 1) Research Question:
##### Does survivors mostly came from the first class during the Titanic incident?

#### 2) Data Sources
##### Primary Source: The data comes from Wikipedia. It was collected after the titanic incident since the survivors were also included in the data.

##### Second Source: From R package called vcdExtra.

##### There are 1309 cases and each of them represents the passenger of the Titanic ship. For this project, I am planning to use survived, class, age and gender variables. 

```{r}
# clean up R environment
rm(list = ls())

# load packages
library(tidyverse)
library(mosaic)
library(mdsr)
library(DataComputing)
library(vcdExtra)
```

```{r}
# Read Primary Data into R
Passenger <- read.csv("full.csv")

# inspect data
head(Passenger)
str(Passenger)
glimpse(Passenger)

# Read second Data Source
data("Titanicp")

# inspect data
head(Titanicp)
str(Titanicp)
glimpse(Titanicp)
```
```{r}
# changing the variables
Passenger$Survived <- as.numeric(Passenger$Survived)
Passenger$Pclass <- as.factor(Passenger$Pclass)
str(Passenger)
```

```{r}
# Creating a new dataframe with some variables that will be used
Survivor <-
  Passenger %>%
  select("PassengerId", "Name", "Pclass", "Survived", "Sex", "Age")

# eliminate incomplete subject
Survivor <- Survivor[!is.na(Survivor$Age),] 

# spreading the Sex variable 
Gender <- 
  Survivor %>%
  group_by(Name, Sex) %>%
  spread(key = Sex, value = Survived)
Gender
```

```{r}
# counting the number of survivors between male and female
Female = sum(Gender$female, na.rm = TRUE)
Female
Male = sum(Gender$male, na.rm = TRUE)
Male

# how many survivors are from class 1, class 2 and class 3
Gender %>%
  filter()
```
There are 233 female survivors and 109 male survivors.

```{r}
# Recall my reserarch question: Does survivors mostly came from the first class during the Titanic incident? 
# We want to see whether passenger's class is significant to the survivors.

RegressModel <- lm(Survived ~ Pclass, data = Survivor)
RegressModel$coefficients 
summary(RegressModel)

treeModel <- rpart(Survived ~ Pclass, data = Survivor)
```

```{r}
# Plot to see the relationship of survivors and classes as well as gender

PassengerList %>%
  ggplot(aes(Age, Survived, color = Pclass)) +
  stat_smooth(method = "glm", family = "binomial", formula = y~x, alpha = .2, aes(fill = Pclass)) +
  facet_grid(.~Sex) +
  theme_bw()
```
From the graph above, it can be seen that the survivors are more among female than male. Between female passengers, the survivors mostly come from class two while between male passengers, survivors mostly come from class one. 




