---
title: "Final Project - Titanic Passengers"
author: "Iffah Mohamad Saufi"
output: html_notebook
---

#### 1) Research Question:
##### Does survivors mostly came from the first class during the Titanic incident?

#### 2) Data Sources
##### The data comes from Wikipedia. It was collected after the titanic incident since the survivors were also included in the data.

##### There are 1309 cases and each of them represents the passenger of the Titanic ship. For this project, I am planning to use survived, class, age and gender variables. 

```{r}
# clean up R environment
rm(list = ls())

# load packages
library(tidyverse)
library(mosaic)
library(mdsr)
library(DataComputing)
```

```{r}
# load Primary Data into R
Passenger <- read.csv("full.csv")

# check data
head(Passenger)
str(Passenger)
glimpse(Passenger)
```

```{r}
# selecting variables that will be used
PassengerList <-
  Passenger %>%
  select("Survived", "Sex", "Pclass", "Age") 

# clean data
PassengerList <- PassengerList[!is.na(PassengerList$Age),] 
```

```{r}
# changing the variables
PassengerList$Pclass <- as.factor(PassengerList$Pclass)
PassengerList$Survived <- as.numeric(PassengerList$Survived)
str(PassengerList)
```

```{r}
# Plot to see the relationship of survivors and classes as well as gender

PassengerList %>%
  ggplot(aes(Age, Survived, color = Pclass)) +
  stat_smooth(method = "glm", family = "binomial", formula = y~x, alpha = .2, aes(fill = Pclass)) +
  facet_grid(.~Sex) +
  theme_bw()
```
From the graph above, it can be seen that the survivors are more among female than male. Between female passengers, the survivors mostly come from class two while between male passengers, survivors mostly come from class one. 

```{r}
# Other Data Source
library(vcdExtra)
data("Titanicp")

head(Titanicp)
str(Titanicp)
glimpse(Titanicp)
```
There are a list of titanic passenger in R package called vcdExtra.

```{r}
# Informative plots
hist(Titanicp$age, xlab = "Age", main = "Age of Titanic Passengers")
```
From the bar graph above, we know that the average of the passenger is around 20-30 years old. 

